---
title: "Assignment02"
output:
  html_document: default
  pdf_document: default
date: "2025-03-28"
---

## Project Overview

A new lifestyle and retail brand has approached you to help them assess data related to buying patterns
in different countries. They are in an expansion mode and would like to present some findings to their
investment board regarding the pros and cons related to some key products that they carry. They are
also interested also in understanding perspectives from travelers in specified countries to enhance their
understanding of how to approach a key segment of their market, vacationers

```{r setup}
#install.packages(c("dplyr", "tidyr", "ggplot2", "plotly", "tm"))
#install.packages("corrplot")
#install.packages(c("dplyr", "tidyr", "ggplot2", "plotly", "tm"))
#install.packages("corrplot")
#install.packages(c("SnowballC","syuzhet","wordcloud","wordcloud2","RColorBrewer"))

library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(tm)
library(corrplot)
library(SnowballC)
library(syuzhet)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)


```



## Part 1

First we will load in traveler data, and create dataframes for three Caribbean countries: Jamaica, The Bahamas and Trinidad and Tobego
```{r}

#Clean up data
rm(list=ls())
options(scipen=9999)

#Load in traveler data
tdata <- read.csv("traveler_reviews.csv")
head(tdata, 20)

#Create subsets of the data for Jamaica, The Bahamas and Trinidad
jam <- subset(tdata, Country == "Jamaica")
head(jam)
baha <- subset(tdata, Country == "The Bahamas")
head(baha)
trini <- subset(tdata, Country == "Trinidad and Tobago")
head(trini)
```
## Wordclouds

# The Bahamas
```{r Bahamas wordcloud}
#Get reviews for wordclouds
baha_rev <- baha$ReviewText

#Convert to corpus and convert to lowercase
btext <- Corpus(VectorSource(baha_rev))
btext <- tm_map(btext, content_transformer(tolower))


#Remove stopwords
btext <- tm_map(btext, removeWords, c(stopwords("english"), "and", "or","hotel"))


#Create wordcloud
btext.dtm <- TermDocumentMatrix(btext)

btext.matrix <- as.matrix(btext.dtm)

bwords <- sort(rowSums(btext.matrix), decreasing=TRUE)

btext.df <- data.frame(word = names(bwords),freq=bwords)
wordcloud2(data=btext.df, size = 0.20, shape="ellipse")

```
# Jamaica 

```{r Jamaica wordcloud}
#Get reviews for wordclouds
jam_rev <- jam$ReviewText

#Convert to corpus and convert to lowercase
jtext <- Corpus(VectorSource(jam_rev))
jtext <- tm_map(jtext, content_transformer(tolower))

#Remove stopwords
jtext <- tm_map(jtext, removeWords, c(stopwords("english"), "and", "or","hotel"))

#Create wordcloud
jtext.dtm <- TermDocumentMatrix(jtext)

jtext.matrix <- as.matrix(jtext.dtm)

jwords <- sort(rowSums(jtext.matrix), decreasing=TRUE)

jtext.df <- data.frame(word = names(jwords),freq=jwords)
wordcloud2(data=jtext.df, size = 0.20, shape="ellipse")
```
# Trinidad
```{r Trinidad}
#Get reviews for wordclouds
trini_rev <- trini$ReviewText
head(trini_rev)

#Convert to corpus and convert to lowercase
ttext <- Corpus(VectorSource(trini_rev))
ttext <- tm_map(ttext, content_transformer(tolower))

#Remove stopwords
ttext <- tm_map(ttext, removeWords, c(stopwords("english"), "and", "or","hotel"))

#Create wordcloud
ttext.dtm <- TermDocumentMatrix(ttext)

ttext.matrix <- as.matrix(ttext.dtm)

twords <- sort(rowSums(ttext.matrix), decreasing=TRUE)

ttext.df <- data.frame(word = names(twords),freq=twords)
wordcloud2(data=ttext.df, size = 0.20, shape="ellipse")
```
## Wordcloud Comparisons
Based on the word clouds, it is apparent that most reviews for Jamaica and Trinidad and Tobego are about the hotel room, since the largest word in their wordclouds is "room". On the other hand, most reviews in the Bahamas discuss "breakfast"

## Sentiment Analysis for each Country

# Bahamas

```{r bahamas sentiment }
#Get sentiments and add to table
bsent <- get_nrc_sentiment(baha$ReviewText)
bvibes <- cbind(baha, bsent)
head(bvibes)

#Determine dominant sentiment
count <- colSums(bsent)
most <- names(count)[which.max(count)]
cat("The most common sentiment in The Bahamas is", most)

```

#Jamaica

```{r jamaica sentiment }
#Get sentiments and add to table
jsent <- get_nrc_sentiment(jam$ReviewText)
jvibes <- cbind(jam, jsent)
head(jvibes)

#Determine dominant sentiment
count <- colSums(jsent)
most <- names(count)[which.max(count)]
cat("The most common sentiment in Jamaica is", most)

```

#Trinidad and Tobego

```{r trinidad sentiment }
#Get sentiments and add to table
tsent <- get_nrc_sentiment(trini$ReviewText)
tvibes <- cbind(trini, tsent)
head(tvibes)

#Determine dominant sentiment
count <- colSums(tsent)
most <- names(count)[which.max(count)]
cat("The most common sentiment in Trinidad and Tobago is", most)

```


#-------------PART 2

#------------- Deleting stored Objects and Importing csv file 


```{r load data}

rm(list=ls())
options(scipen=9999)
sales.df <- read.csv("sales_data03.csv", stringsAsFactors = FALSE)
head(sales.df)

```
#-------------- Deleting rows entirely consisting of "", NA or "k"s. 

```{r}
#Step 1 - Removing empty rows
noNARow <- sales.df[rowSums(is.na(sales.df[, -1]) | sales.df[, -1] == "" | sales.df[, -1] == "k") != (ncol(sales.df) - 1), ]

```


#---------------- Checking how dirty the Region feature is and extracting only the valid regions

```{r}

uniqueValues <- unique(noNARow[,"Region"])
print(uniqueValues)

validPlaces <- c("Sub-Saharan Africa", "Asia" , "Europe" , "Australia and Oceania", "Central America and the Caribbean",
                 "Middle East and North Africa", "North America")

print(validPlaces)
```


#------------------- Removing extra places at the end of countries eg. "Saint Kitts and Nevis " and also getting all the countries

```{r}

noNARow$Country <- trimws(noNARow$Country)
uniqueValues <- unique(noNARow[,"Country"])
print(uniqueValues)
```


#------------------- As all entries under Countries are valid we can extract the missing values or the region by using domain knowledge
#here we assign each country to their region


```{r}
countryRegionMap <- c(
  "Lesotho" = "Sub-Saharan Africa", "Indonesia" = "Asia", "Norway" = "Europe",
  "Latvia" = "Europe", "Uzbekistan" = "Asia", "Papua New Guinea" = "Australia and Oceania",
  "Serbia" = "Europe", "Panama" = "Central America and the Caribbean", "Jamaica" = "Central America and the Caribbean",
  "Brunei" = "Asia", "Senegal" = "Sub-Saharan Africa", "San Marino" = "Europe",
  "Vanuatu" = "Australia and Oceania", "East Timor" = "Asia", "Slovenia" = "Europe",
  "Slovakia" = "Europe", "Croatia" = "Europe", "Niger" = "Sub-Saharan Africa",
  "Austria" = "Europe", "Tanzania" = "Sub-Saharan Africa", "Eritrea" = "Sub-Saharan Africa",
  "Luxembourg" = "Europe", "Sao Tome and Principe" = "Sub-Saharan Africa", "Mali" = "Sub-Saharan Africa",
  "Kuwait" = "Middle East and North Africa", "Tuvalu" = "Australia and Oceania", "Saudi Arabia" = "Middle East and North Africa",
  "Cambodia" = "Asia", "Kyrgyzstan" = "Asia", "Central African Republic" = "Sub-Saharan Africa",
  "Denmark" = "Europe", "Madagascar" = "Sub-Saharan Africa", "Hungary" = "Europe",
  "United Kingdom" = "Europe", "Ethiopia" = "Sub-Saharan Africa", "South Korea" = "Asia",
  "Mexico" = "North America", "United Arab Emirates" = "Middle East and North Africa",
  "Sweden" = "Europe", "Syria" = "Middle East and North Africa", "Mauritius" = "Sub-Saharan Africa",
  "Honduras" = "Central America and the Caribbean", "Comoros" = "Sub-Saharan Africa",
  "Mozambique" = "Sub-Saharan Africa", "Qatar" = "Middle East and North Africa",
  "Vietnam" = "Asia", "Sri Lanka" = "Asia", "Bosnia and Herzegovina" = "Europe",
  "Costa Rica" = "Central America and the Caribbean", "Cape Verde" = "Sub-Saharan Africa",
  "Japan" = "Asia", "Italy" = "Europe", "Samoa" = "Australia and Oceania",
  "Guatemala" = "Central America and the Caribbean", "South Africa" = "Sub-Saharan Africa",
  "Barbados" = "Central America and the Caribbean", "Portugal" = "Europe",
  "Swaziland" = "Sub-Saharan Africa", "Mauritania" = "Sub-Saharan Africa",
  "New Zealand" = "Australia and Oceania", "Belize" = "Central America and the Caribbean",
  "Tonga" = "Australia and Oceania", "Georgia" = "Europe", "Taiwan" = "Asia",
  "Bhutan" = "Asia", "Kenya" = "Sub-Saharan Africa", "Algeria" = "Middle East and North Africa",
  "Kiribati" = "Australia and Oceania", "Palau" = "Australia and Oceania",
  "Sierra Leone" = "Sub-Saharan Africa", "Bangladesh" = "Asia", "Thailand" = "Asia",
  "Yemen" = "Middle East and North Africa", "Afghanistan" = "Asia",
  "Trinidad and Tobago" = "Central America and the Caribbean",
  "United States of America" = "North America", "Germany" = "Europe",
  "Andorra" = "Europe", "Turkey" = "Asia", "Singapore" = "Asia",
  "Switzerland" = "Europe", "France" = "Europe", "Uganda" = "Sub-Saharan Africa",
  "Liberia" = "Sub-Saharan Africa", "Kosovo" = "Europe", "Macedonia" = "Europe",
  "Equatorial Guinea" = "Sub-Saharan Africa", "Solomon Islands" = "Australia and Oceania",
  "Turkmenistan" = "Asia", "Belarus" = "Europe", "Netherlands" = "Europe",
  "Romania" = "Europe", "Russia" = "Europe", "Togo" = "Sub-Saharan Africa",
  "Lebanon" = "Middle East and North Africa", "Bulgaria" = "Europe",
  "Cyprus" = "Europe", "Australia" = "Australia and Oceania", "Monaco" = "Europe",
  "Malawi" = "Sub-Saharan Africa", "Saint Lucia" = "Central America and the Caribbean",
  "Montenegro" = "Europe", "Sudan" = "Sub-Saharan Africa", "Nauru" = "Australia and Oceania",
  "Zambia" = "Sub-Saharan Africa", "Kazakhstan" = "Asia", "Oman" = "Middle East and North Africa",
  "Vatican City" = "Europe", "Chad" = "Sub-Saharan Africa", "Armenia" = "Asia",
  "Belgium" = "Europe", "Antigua and Barbuda" = "Central America and the Caribbean",
  "Myanmar" = "Asia", "Iraq" = "Middle East and North Africa",
  "Burundi" = "Sub-Saharan Africa", "Finland" = "Europe", "Tajikistan" = "Asia",
  "Cuba" = "Central America and the Caribbean", "Dominica" = "Central America and the Caribbean",
  "Czech Republic" = "Europe", "Poland" = "Europe", "Malaysia" = "Asia",
  "Greenland" = "North America", "Benin" = "Sub-Saharan Africa", "Azerbaijan" = "Asia",
  "Iran" = "Middle East and North Africa", "Mongolia" = "Asia",
  "Lithuania" = "Europe", "Guinea" = "Sub-Saharan Africa",
  "Estonia" = "Europe", "Tunisia" = "Middle East and North Africa",
  "Saint Vincent and the Grenadines" = "Central America and the Caribbean",
  "North Korea" = "Asia", "Malta" = "Europe", "Liechtenstein" = "Europe",
  "The Bahamas" = "Central America and the Caribbean", "Ukraine" = "Europe",
  "Federated States of Micronesia" = "Australia and Oceania",
  "South Sudan" = "Sub-Saharan Africa", "Moldova" = "Europe",
  "Philippines" = "Asia", "The Gambia" = "Sub-Saharan Africa",
  "Israel" = "Middle East and North Africa", "Burkina Faso" = "Sub-Saharan Africa",
  "Djibouti" = "Sub-Saharan Africa", "Bahrain" = "Middle East and North Africa",
  "Botswana" = "Sub-Saharan Africa", "Haiti" = "Central America and the Caribbean",
  "Somalia" = "Sub-Saharan Africa", "Namibia" = "Sub-Saharan Africa",
  "Marshall Islands" = "Australia and Oceania", "Greece" = "Europe",
  "Spain" = "Europe", "Saint Kitts and Nevis" = "Central America and the Caribbean",
  "Seychelles" = "Sub-Saharan Africa", "Dominican Republic" = "Central America and the Caribbean",
  "Albania" = "Europe", "Rwanda" = "Sub-Saharan Africa", "Gabon" = "Sub-Saharan Africa",
  "Canada" = "North America", "Ghana" = "Sub-Saharan Africa",
  "Morocco" = "Middle East and North Africa", "Guinea-Bissau" = "Sub-Saharan Africa",
  "Ireland" = "Europe", "India" = "Asia", "Democratic Republic of the Congo" = "Sub-Saharan Africa",
  "Iceland" = "Europe", "Pakistan" = "Asia", "Laos" = "Asia",
  "Cameroon" = "Sub-Saharan Africa", "Cote d'Ivoire" = "Sub-Saharan Africa",
  "Grenada" = "Central America and the Caribbean", "Maldives" = "Asia",
  "Nepal" = "Asia", "Fiji" = "Australia and Oceania", "Nicaragua" = "Central America and the Caribbean",
  "Jordan" = "Middle East and North Africa", "El Salvador" = "Central America and the Caribbean",
  "Egypt" = "Middle East and North Africa", "Republic of the Congo" = "Sub-Saharan Africa",
  "Angola" = "Sub-Saharan Africa", "Zimbabwe" = "Sub-Saharan Africa", 
  "Nigeria" = "Sub-Saharan Africa"
)
```



#---------------------------- Invalid Region entries are replaced with the correct Regions based on their country

```{r}
fixedCountries <- noNARow
fixedCountries$Region <- ifelse(noNARow$Region %in% validPlaces, noNARow$Region, countryRegionMap[noNARow$Country])

uniqueValues <- unique(fixedCountries[,"Region"])
print(uniqueValues)
```



#------------------------- X is reviewed but we decided to view "None" as the option "other"

```{r}
#Note - None is one of the options
uniqueValues <- unique(fixedCountries[,"X"])
print(uniqueValues)
```

#------------------------- Sales.Channel is checked on what dirty values it has.

```{r}
#Note - Yes is one of the options
uniqueValues <- unique(fixedCountries[,"Sales.Channel"])
print(uniqueValues)
fixedCountries[fixedCountries$Sales.Channel == "", ]
```

```{r}
# Print the unique values in the Sales.Channel column to check
uniqueValue <- unique(fixedCountries[,"Sales.Channel"])
print(uniqueValue)
```


#----------------------- Using predictions we predict if the Sales.Channel based on the country and item type


```{r}


fixedCountries$Sales.Channel <- tolower(fixedCountries$Sales.Channel)

fixedCountries$Sales.Channel <- ifelse(fixedCountries$Sales.Channel %in% c("yes", ""), NA, fixedCountries$Sales.Channel)

fixedCountries$Sales.Channel <- factor(fixedCountries$Sales.Channel, levels = c("offline", "online"))

print(table(fixedCountries$Sales.Channel, useNA = "always"))

train_data <- fixedCountries %>% filter(!is.na(Sales.Channel))

if (length(unique(train_data$Sales.Channel)) > 1) {
    model <- glm(Sales.Channel ~ Country + X, data = train_data, family = binomial, na.action = na.exclude)

    missing_indices <- which(is.na(fixedCountries$Sales.Channel))
    fixedCountries$PredictedProb <- predict(model, newdata = fixedCountries, type = "response")

    fixedCountries$Sales.Channel[missing_indices] <- ifelse(fixedCountries$PredictedProb[missing_indices] > 0.5, "online", "offline")

    fixedCountries$Sales.Channel <- factor(fixedCountries$Sales.Channel, levels = c("offline", "online"))

    fixedCountries <- fixedCountries %>% select(-PredictedProb)
} else {
    warning("Not enough unique levels in Sales.Channel to fit the model. Check data cleaning steps!")
}

print(table(fixedCountries$Sales.Channel, useNA = "always"))

```

#------------------------- Order.Date and Ship.Date are compared to check for errors

```{r}

fixedCountries$Order.Date <- as.Date(fixedCountries$Order.Date, format="%m/%d/%Y")
fixedCountries$Ship.Date <- as.Date(fixedCountries$Ship.Date, format="%m/%d/%Y")

all(fixedCountries$Order.Date <= fixedCountries$Ship.Date)


#invalid_rows <- fixedCountries[fixedCountries$Order.Date > fixedCountries$Ship.Date, ]

# Print the problematic rows
#print(invalid_rows)

```


#-------------------------- Region

```{r}
summary(fixedCountries$Region)
uniqueValues <- unique(fixedCountries[,"Region"])
print(uniqueValues)
```




```{r}

product_counts <- fixedCountries %>%
  count(Region) 

ggplot(product_counts, aes(x = reorder(Region, -n), y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
    title = "Frequency of Region",
    x = "Count",
    y = "Region"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-labels

```



#----------------------------------------------- Country

```{r}
summary(fixedCountries$Country)
uniqueValues <- unique(fixedCountries[,"Country"])
print(uniqueValues)
```


```{r}

top_countries <- fixedCountries %>%
  count(Country) %>%
  top_n(10, n) 

ggplot(top_countries, aes(x = reorder(Country, -n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Top 10 Countries by Sales Frequency", x = "Country", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```







#--------------------------------- X

```{r}
summary(fixedCountries$X)
uniqueValues <- unique(fixedCountries[,"X"])
print(uniqueValues)
```

#-------------------------- It was noticed that Unit.Cost only relates to one type of item. Using this we are able to look at the price
# to figure out the X eg 117.11 is only related to cereal

```{r}
cost_to_product <- fixedCountries %>%
  filter(X != "None" & !is.na(X) & X != "") %>%
  group_by(Unit.Cost) %>%
  summarize(
    potential_product = first(X), 
    product_count = n_distinct(X)  
  ) %>%
  filter(product_count == 1) 

fixedCountries <- fixedCountries %>%
  left_join(cost_to_product, by = "Unit.Cost") %>%
  mutate(
    X = case_when(
      X == "None" & !is.na(potential_product) ~ potential_product,
      TRUE ~ X
    )
  ) %>%
  select(-potential_product, -product_count) 


fixedCountries <- fixedCountries %>%
  mutate(X = ifelse(X == "None", "Other", X))
```


```{r}
summary(fixedCountries$X)
uniqueValues <- unique(fixedCountries[,"X"])
print(uniqueValues)
```



```{r}


product_counts <- fixedCountries %>%
  count(X) 

ggplot(product_counts, aes(x = reorder(X, -n), y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
    title = "Frequency of Product Types (X)",
    x = "Product Type",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-labels


```




#---------------------------------- Sales.Channel


```{r}
summary(fixedCountries$Sales.Channel)
uniqueValues <- unique(fixedCountries[,"Sales.Channel"])
print(uniqueValues)


```
```{r}

channel_counts <- fixedCountries %>%
  count(Sales.Channel) %>%
  arrange(desc(n))  # Sort by frequency (descending)

ggplot(channel_counts, aes(x = Sales.Channel, y = n)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Frequency of Sales by Channel", x = "Sales Channel", y = "Count") +
  theme_minimal()
```


#------------------------------- Order.Priority

```{r}
summary(fixedCountries$Order.Priority)
uniqueValues <- unique(fixedCountries[,"Order.Priority"])
print(uniqueValues)
```

```{r}
channel_counts <- fixedCountries %>%
  count(Order.Priority) %>%
  arrange(desc(n))  # Sort by frequency (descending)

ggplot(channel_counts, aes(x = Order.Priority, y = n)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Frequency of Sales by Order.Priority", x = "Order.Priority", y = "Count") +
  theme_minimal()
```

#---------------------------------- Order.Date

```{r}
summary(fixedCountries$Order.Date)


```

```{r}
ggplot(fixedCountries, aes(x = Order.Date)) +
  geom_line(stat = "count", color = "red", linewidth = .05) +  # Count orders per date
  labs(
    title = "Orders Over Time (Check for Gaps/Spikes)",
    x = "Order Date",
    y = "Number of Orders"
  ) +
  theme_minimal()
```


#---------------------------------- Ship.Date


```{r}
summary(fixedCountries$Ship.Date)

```


```{r}
ggplot(fixedCountries, aes(x = Ship.Date)) +
  geom_line(stat = "count", color = "red", linewidth = .05) +  # Count orders per date
  labs(
    title = "Ship.Date (Check for Gaps/Spikes)",
    x = "Ship Date",
    y = "Number of Orders"
  ) +
  theme_minimal()

channel_counts <- fixedCountries %>%
  count(Ship.Date) %>%
  arrange(desc(n))  # Sort by frequency (descending)

ggplot(channel_counts, aes(x = Ship.Date, y = n)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Frequency of Sales by Ship.Date", x = "Ship.Date", y = "Count") +
  theme_minimal()


```


#--------------------------------- Order.ID

```{r}
fixedCountries$Order.ID <- as.integer(fixedCountries$Order.ID)
summary(fixedCountries$Order.ID)
```

#--------------------------------- Units.Sold

```{r}
fixedCountries$Units.Sold <- as.integer(fixedCountries$Units.Sold)
summary(fixedCountries$Units.Sold)
```



```{r}
ggplot(fixedCountries, aes(x = Units.Sold)) +
  geom_histogram(bins = 30, fill = "purple", color = "white") +
  labs(title = "Distribution of Units Sold", x = "Units Sold", y = "Frequency") +
  theme_minimal()
```

#---------------------------------- Unit.Cost


```{r}
fixedCountries$Unit.Cost <- as.numeric(fixedCountries$Unit.Cost)
summary(fixedCountries$Unit.Cost)
```


```{r}
# Boxplot for Unit.Price
ggplot(fixedCountries, aes(y = Unit.Price)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Boxplot of Unit Price")

# Scatter Plot: Unit.Price vs. Unit.Cost (should generally be above the diagonal)
ggplot(fixedCountries, aes(x = Unit.Cost, y = Unit.Price)) +
  geom_point(color = "blue") +
  geom_abline(slope = 1, linetype = "dashed", color = "red") +  # Profit=0% line
  labs(title = "Unit Price vs. Unit Cost (Check for Pricing Errors)")
```


#---------------------------------- Total.Cost


```{r}
fixedCountries$Total.Cost <- as.numeric(fixedCountries$Total.Cost)
summary(fixedCountries$Total.Cost)
```


```{r}
# Boxplot
ggplot(fixedCountries, aes(y = Total.Cost)) +
  geom_boxplot(fill = "pink") +
  labs(title = "Boxplot of Total Cost")

# Scatter Plot: Total.Cost vs. Total.Revenue
ggplot(fixedCountries, aes(x = Total.Cost, y = Total.Revenue)) +
  geom_point(color = "darkgreen") +
  geom_abline(slope = .05, linetype = "dashed", color = "red") +  # For every $1 of Total.Cost, you expect $0.05 of Total.Revenue to break even.
  labs(title = "Total Cost vs. Total Revenue (Should Mostly Be Below Diagonal)")
```

#------------------------- Total.Revenue


```{r}
fixedCountries$Total.Revenue <- as.numeric(fixedCountries$Total.Revenue)
summary(fixedCountries$Total.Revenue)
```


#--- Using fixedCountries$Total.Profit + fixedCountries$Total.Cost = fixedCountries$Total.Revenue to replace the NA's


```{r}
fixedCountries$Total.Profit <- as.numeric(fixedCountries$Total.Profit)

fixedCountries$Total.Revenue <- ifelse(is.na(fixedCountries$Total.Revenue) & !is.na(fixedCountries$Total.Cost) & !is.na(fixedCountries$Total.Profit),
                          fixedCountries$Total.Profit + fixedCountries$Total.Cost,
                          fixedCountries$Total.Revenue)

summary(fixedCountries$Total.Revenue)
```


```{r}
# Histogram
ggplot(fixedCountries, aes(x = Total.Revenue)) +
  geom_histogram(bins = 30, fill = "steelblue") +
  labs(title = "Distribution of Total Revenue")

# Time Series (if Order.Date exists)
fixedCountries$Order.Date <- as.Date(fixedCountries$Order.Date)  # Convert to date
ggplot(fixedCountries, aes(x = Order.Date, y = Total.Revenue)) +
  geom_line(color = "darkred") +
  labs(title = "Total Revenue Over Time")

revenue_by_product <- fixedCountries %>%
  group_by(X) %>%
  summarise(Total_Revenue = sum(Total.Revenue, na.rm = TRUE)) %>%
  arrange(desc(Total_Revenue))  # Sort by highest revenue

ggplot(revenue_by_product, aes(x = reorder(X, -Total_Revenue), y = Total_Revenue)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
    title = "Total Revenue by Product Type (X)",
    x = "Product Type",
    y = "Total Revenue"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-labels
```

#------------------------- Unit.Price




```{r}
fixedCountries$Unit.Price <- as.numeric(fixedCountries$Unit.Price)
summary(fixedCountries$Unit.Price)
```

#--- Using fixedCountries$Total.Revenue / fixedCountries$Units.Sold = Unit.Price to replace the NA's

```{r}
fixedCountries$Unit.Price <- ifelse(
  is.na(fixedCountries$Unit.Price) & 
  !is.na(fixedCountries$Total.Revenue) & 
  !is.na(fixedCountries$Units.Sold),
  fixedCountries$Total.Revenue / fixedCountries$Units.Sold,  # Calculate Unit.Price
  fixedCountries$Unit.Price  # Keep original value if not NA
)

summary(fixedCountries$Unit.Price)
```


```{r}
# Boxplot for Unit.Price
ggplot(fixedCountries, aes(y = Unit.Price)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Boxplot of Unit Price")

# Scatter Plot: Unit.Price vs. Unit.Cost (should generally be above the diagonal)
ggplot(fixedCountries, aes(x = Unit.Cost, y = Unit.Price)) +
  geom_point(color = "blue") +
  geom_abline(slope = 1, linetype = "dashed", color = "red") +  # Profit=0 line
  labs(title = "Unit Price vs. Unit Cost (Check for Pricing Errors)")
```


#------------------------- Total.Profit

```{r}
fixedCountries$Total.Profit <- as.numeric(fixedCountries$Total.Profit)
summary(fixedCountries$Total.Profit)
```

#--- Using fixedCountries$Total.Revenue - fixedCountries$Total.Cost = Total.Profit to replace the NA's

```{r}
fixedCountries$Total.Profit <- ifelse(is.na(fixedCountries$Total.Profit) & !is.na(fixedCountries$Total.Revenue) & !is.na(fixedCountries$Total.Cost),
                          fixedCountries$Total.Revenue - fixedCountries$Total.Cost,
                          fixedCountries$Total.Profit)

summary(fixedCountries$Total.Profit)

```


```{r}

# Boxplot (identify extreme values)
ggplot(fixedCountries, aes(y = Total.Profit)) +
  geom_boxplot(fill = "skyblue") +
  labs(title = "Boxplot of Total Profit (Outlier Detection)")

# Histogram (check distribution)
ggplot(fixedCountries, aes(x = Total.Profit)) +
  geom_histogram(bins = 30, fill = "purple") +
  labs(title = "Distribution of Total Profit")

# Density plot (smooth distribution)
ggplot(fixedCountries, aes(x = Total.Profit)) +
  geom_density(fill = "orange", alpha = 0.7) +
  labs(title = "Density of Total Profit")
```


#-------------------------


```{r}
cleanData.df = fixedCountries
```



#------------------------ PART 3

## Analysis


# What is the average amount spent on Personal Care vs. Cosmetics in each region?

```{r a}

#Clean up data
#rm(list=ls())
#options(scipen=9999)

sales.df = cleanData.df


#Convert Revenue to numeric
sales.df$Total.Revenue <- as.numeric(sales.df$Total.Revenue)

#Create dataframes for Personal Care and Cosmetics
caredata <- subset(sales.df, X == "Personal Care")


cosmeticsdata <- subset(sales.df, X == "Cosmetics")


careavg <- mean(caredata$Total.Revenue, na.rm = TRUE)
cosmeticsavg <- mean(cosmeticsdata$Total.Revenue, na.rm = TRUE)

cat("The average money spent on Personal care is", careavg, "while the average spent on Cosmetics is", cosmeticsavg)
```
# In which country do people use the most cosmetics? 

```{r b}
#Convert Revenue to numeric
cosmeticsdata$Units.Sold <- as.numeric(cosmeticsdata$Units.Sold)

#Get sums the units sold in each country
cosmeticsum <- aggregate(Units.Sold ~ Country, data = cosmeticsdata, FUN = sum, na.rm = TRUE)

#Gets the data for the country with the maximum units sold
cosmax <- cosmeticsum[which.max(cosmeticsum$Units.Sold), ]
cat("People use the most cosmetics in", cosmax$Country)
```
# In 2016, which product had the highest and lowest quantity sold across each region?

```{r c}
#Set datatypes
sales.df$Order.Date <- as.Date(sales.df$Order.Date, format = "%d/%m/%Y")
sales.df$Units.Sold <- as.numeric(sales.df$Units.Sold)

#Create subset of data from 2016
data2016 <- subset(sales.df, format(Order.Date, "%Y") == "2016")

#Find total per region and per product
total <- aggregate(Units.Sold ~ Region + X, data = data2016, sum, na.rm = TRUE)

#Split data by regions
byregions <- split(total, total$Region)

#Create a dataframe with the maximums and minimums found from byregions

hilo <- lapply(byregions, function(df) {
  data.frame(
    Region = unique(df$Region),
    Highest.Sold = df$X[which.max(df$Units.Sold)],
    Lowest.Sold = df$X[which.min(df$Units.Sold)]
  )
})

hilo <- do.call(rbind, hilo)
print(hilo)

```
# What is the average profit for each item by region? 

```{r d}
#Set datatypes
sales.df$Total.Profit <- as.numeric(sales.df$Total.Profit)

aggregate(Total.Profit ~ Region + X, data = sales.df, FUN = mean, na.rm = TRUE)
```
# In which of the seasons (Spring, Summer, Autumn, Winter) do people spend the most on:  
Cosmetics
Personal Care
```{r e}
#Note that we will use the same seasons for all regions, and assume typical Northern Hemisphere Seasons where, generally
#Winter is December, January and February
#Spring is March, April and May
#Summer is June, July and August
#Autumn is September, October, November

#Set datatypes
sales.df$Month <- as.numeric(format(sales.df$Order.Date, "%m"))

#Create bins labelled as seasons, and assign months to seasons
sales.df$Season <- cut(sales.df$Month,
                       breaks = c(0, 2, 5, 8, 11, 12),
                       labels = c("Winter", "Spring", "Summer", "Autumn", "Winter"),
                       include.lowest = TRUE, right = TRUE)

#Create an aggregate of revenue by season
seasonsales <- aggregate(Total.Revenue ~ Season + X, data = sales.df[sales.df$X %in% c("Cosmetics", "Personal Care"), ], sum, na.rm = TRUE)


#Set datatypes
seasonsales$Season <- as.character(seasonsales$Season)

#Get maximums
maxcos <- seasonsales$Season[which.max(seasonsales$Total.Revenue[seasonsales$X == "Cosmetics"])]
maxcare <- seasonsales$Season[which.max(seasonsales$Total.Revenue[seasonsales$X == "Personal Care"])]

cat("Highest spending on cosmetics was in", maxcos, "and highest spending on personal care was in", maxcare)
```
# Find the monthly average units sold for Personal Care, Cosmetics and Beverages and plot a series graph for ONE year . Discuss the resulting visualization.
```{r f}
#Remove days
data2016$YearMonth <- format(data2016$Order.Date, "%Y-%m")

#Create subset with only the needed data
data2016filter<- data2016[data2016$X %in% c("Personal Care", "Cosmetics", "Beverages"), ]

#Find monthly average
monthlyavg <- aggregate(Units.Sold ~ YearMonth + X, data = data2016filter, FUN = mean, na.rm = TRUE)

#Plot line graph
ggplot(monthlyavg, aes(x = YearMonth, y = Units.Sold, color = X, group = X)) + geom_line() + labs(title = "Monthly Average Units Sold", x = "Month", y = "Units Sold") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Determine correlation between Total.Profit and the other features (include visualizations) in the dataset.

Units Sold and Total Profit are x correlated
Unit Price and Total Profit are x correlated
Unit Cost and Total Profit are x correlated
Total Revenue and Total Profit are x correlated
Total Cost and Total Profit are x correlated
```{r g}
#Set datatypes
salesg.df <- sales.df %>% select(Total.Profit, Units.Sold, Unit.Price, Unit.Cost, Total.Revenue, Total.Cost) %>%mutate(across(everything(), as.numeric))

#Create correlation matrix
num_cols <- sapply(salesg.df, is.numeric)
cor_matrix <- cor(salesg.df[, num_cols], use = "complete.obs")
profit_corr <- cor_matrix["Total.Profit", , drop = FALSE]

#Visualize correlation with Profit
corrplot(profit_corr, method = "color")

#Check correlations
cor(salesg.df$Units.Sold,sales.df$Total.Profit)
cor(salesg.df$Unit.Price,sales.df$Total.Profit)
cor(salesg.df$Unit.Cost,sales.df$Total.Profit)
cor(salesg.df$Total.Revenue,sales.df$Total.Profit)
cor(salesg.df$Total.Cost,sales.df$Total.Profit)
```





# Based on correlation from part (g), build a predictive model that can be used to predict Total.Profit. Determine the accuracy of the model and show the result of a sample set of predictions for data not included in the given dataset.
```{r h}
#Split data into training and testing
set.seed(101)
train <- (sample(1:nrow(sales.df),nrow(sales.df)*0.8))
training <- sales.df[train,]
testing <- sales.df[-train,]

#Model
result <- result <- lm(Total.Profit ~ Units.Sold + Unit.Price + Unit.Cost + Total.Revenue + Total.Cost, data = sales.df)
summary(result)

#Accuracy
testing$predictions <- as.integer(predict(result,testing))
sum(ifelse(abs(testing$Total.Profit-testing$predictions)<=5,1,0))/nrow(testing)*100
```


# Explain how this model may help the investors make better decisions.

Investors could use this model to predict profits. This is especially useful when deciding how much of a product should be purchased, and how that may translate to future profits


