---
title: "Assignment02"
output: html_document
date: "2025-03-28"
---

## Project Overview

A new lifestyle and retail brand has approached you to help them assess data related to buying patterns
in different countries. They are in an expansion mode and would like to present some findings to their
investment board regarding the pros and cons related to some key products that they carry. They are
also interested also in understanding perspectives from travelers in specified countries to enhance their
understanding of how to approach a key segment of their market, vacationers

```{r setup}
#install.packages(c("dplyr", "tidyr", "ggplot2", "plotly", "tm"))
#install.packages("corrplot")

library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(tm)
library(corrplot)


```

#------------- Deleting stored Objects and Importing csv file 


```{r load data}

rm(list=ls())
options(scipen=9999)

tdata <- read.csv("traveler_reviews.csv")
head(tdata)
```



```{r}
sales.df <- read.csv("sales_data03.csv", stringsAsFactors = FALSE)
```

#-------------- Deleting rows entirely consisting of "", NA or "k"s. 

```{r}
#Step 1 - Removing empty rows
noNARow <- sales.df[rowSums(is.na(sales.df[, -1]) | sales.df[, -1] == "" | sales.df[, -1] == "k") != (ncol(sales.df) - 1), ]

```


#---------------- Checking how dirty the Region feature is and extracting only the valid regions

```{r}

uniqueValues <- unique(noNARow[,"Region"])
print(uniqueValues)

validPlaces <- c("Sub-Saharan Africa", "Asia" , "Europe" , "Australia and Oceania", "Central America and the Caribbean",
                 "Middle East and North Africa", "North America")

print(validPlaces)
```


#------------------- Removing extra places at the end of countries eg. "Saint Kitts and Nevis " and also getting all the countries

```{r}

noNARow$Country <- trimws(noNARow$Country)
uniqueValues <- unique(noNARow[,"Country"])
print(uniqueValues)
```


#------------------- As all entries under Countries are valid we can extract the missing values or the region by using domain knowledge
#here we assign each country to their region


```{r}
countryRegionMap <- c(
  "Lesotho" = "Sub-Saharan Africa", "Indonesia" = "Asia", "Norway" = "Europe",
  "Latvia" = "Europe", "Uzbekistan" = "Asia", "Papua New Guinea" = "Australia and Oceania",
  "Serbia" = "Europe", "Panama" = "Central America and the Caribbean", "Jamaica" = "Central America and the Caribbean",
  "Brunei" = "Asia", "Senegal" = "Sub-Saharan Africa", "San Marino" = "Europe",
  "Vanuatu" = "Australia and Oceania", "East Timor" = "Asia", "Slovenia" = "Europe",
  "Slovakia" = "Europe", "Croatia" = "Europe", "Niger" = "Sub-Saharan Africa",
  "Austria" = "Europe", "Tanzania" = "Sub-Saharan Africa", "Eritrea" = "Sub-Saharan Africa",
  "Luxembourg" = "Europe", "Sao Tome and Principe" = "Sub-Saharan Africa", "Mali" = "Sub-Saharan Africa",
  "Kuwait" = "Middle East and North Africa", "Tuvalu" = "Australia and Oceania", "Saudi Arabia" = "Middle East and North Africa",
  "Cambodia" = "Asia", "Kyrgyzstan" = "Asia", "Central African Republic" = "Sub-Saharan Africa",
  "Denmark" = "Europe", "Madagascar" = "Sub-Saharan Africa", "Hungary" = "Europe",
  "United Kingdom" = "Europe", "Ethiopia" = "Sub-Saharan Africa", "South Korea" = "Asia",
  "Mexico" = "North America", "United Arab Emirates" = "Middle East and North Africa",
  "Sweden" = "Europe", "Syria" = "Middle East and North Africa", "Mauritius" = "Sub-Saharan Africa",
  "Honduras" = "Central America and the Caribbean", "Comoros" = "Sub-Saharan Africa",
  "Mozambique" = "Sub-Saharan Africa", "Qatar" = "Middle East and North Africa",
  "Vietnam" = "Asia", "Sri Lanka" = "Asia", "Bosnia and Herzegovina" = "Europe",
  "Costa Rica" = "Central America and the Caribbean", "Cape Verde" = "Sub-Saharan Africa",
  "Japan" = "Asia", "Italy" = "Europe", "Samoa" = "Australia and Oceania",
  "Guatemala" = "Central America and the Caribbean", "South Africa" = "Sub-Saharan Africa",
  "Barbados" = "Central America and the Caribbean", "Portugal" = "Europe",
  "Swaziland" = "Sub-Saharan Africa", "Mauritania" = "Sub-Saharan Africa",
  "New Zealand" = "Australia and Oceania", "Belize" = "Central America and the Caribbean",
  "Tonga" = "Australia and Oceania", "Georgia" = "Europe", "Taiwan" = "Asia",
  "Bhutan" = "Asia", "Kenya" = "Sub-Saharan Africa", "Algeria" = "Middle East and North Africa",
  "Kiribati" = "Australia and Oceania", "Palau" = "Australia and Oceania",
  "Sierra Leone" = "Sub-Saharan Africa", "Bangladesh" = "Asia", "Thailand" = "Asia",
  "Yemen" = "Middle East and North Africa", "Afghanistan" = "Asia",
  "Trinidad and Tobago" = "Central America and the Caribbean",
  "United States of America" = "North America", "Germany" = "Europe",
  "Andorra" = "Europe", "Turkey" = "Asia", "Singapore" = "Asia",
  "Switzerland" = "Europe", "France" = "Europe", "Uganda" = "Sub-Saharan Africa",
  "Liberia" = "Sub-Saharan Africa", "Kosovo" = "Europe", "Macedonia" = "Europe",
  "Equatorial Guinea" = "Sub-Saharan Africa", "Solomon Islands" = "Australia and Oceania",
  "Turkmenistan" = "Asia", "Belarus" = "Europe", "Netherlands" = "Europe",
  "Romania" = "Europe", "Russia" = "Europe", "Togo" = "Sub-Saharan Africa",
  "Lebanon" = "Middle East and North Africa", "Bulgaria" = "Europe",
  "Cyprus" = "Europe", "Australia" = "Australia and Oceania", "Monaco" = "Europe",
  "Malawi" = "Sub-Saharan Africa", "Saint Lucia" = "Central America and the Caribbean",
  "Montenegro" = "Europe", "Sudan" = "Sub-Saharan Africa", "Nauru" = "Australia and Oceania",
  "Zambia" = "Sub-Saharan Africa", "Kazakhstan" = "Asia", "Oman" = "Middle East and North Africa",
  "Vatican City" = "Europe", "Chad" = "Sub-Saharan Africa", "Armenia" = "Asia",
  "Belgium" = "Europe", "Antigua and Barbuda" = "Central America and the Caribbean",
  "Myanmar" = "Asia", "Iraq" = "Middle East and North Africa",
  "Burundi" = "Sub-Saharan Africa", "Finland" = "Europe", "Tajikistan" = "Asia",
  "Cuba" = "Central America and the Caribbean", "Dominica" = "Central America and the Caribbean",
  "Czech Republic" = "Europe", "Poland" = "Europe", "Malaysia" = "Asia",
  "Greenland" = "North America", "Benin" = "Sub-Saharan Africa", "Azerbaijan" = "Asia",
  "Iran" = "Middle East and North Africa", "Mongolia" = "Asia",
  "Lithuania" = "Europe", "Guinea" = "Sub-Saharan Africa",
  "Estonia" = "Europe", "Tunisia" = "Middle East and North Africa",
  "Saint Vincent and the Grenadines" = "Central America and the Caribbean",
  "North Korea" = "Asia", "Malta" = "Europe", "Liechtenstein" = "Europe",
  "The Bahamas" = "Central America and the Caribbean", "Ukraine" = "Europe",
  "Federated States of Micronesia" = "Australia and Oceania",
  "South Sudan" = "Sub-Saharan Africa", "Moldova" = "Europe",
  "Philippines" = "Asia", "The Gambia" = "Sub-Saharan Africa",
  "Israel" = "Middle East and North Africa", "Burkina Faso" = "Sub-Saharan Africa",
  "Djibouti" = "Sub-Saharan Africa", "Bahrain" = "Middle East and North Africa",
  "Botswana" = "Sub-Saharan Africa", "Haiti" = "Central America and the Caribbean",
  "Somalia" = "Sub-Saharan Africa", "Namibia" = "Sub-Saharan Africa",
  "Marshall Islands" = "Australia and Oceania", "Greece" = "Europe",
  "Spain" = "Europe", "Saint Kitts and Nevis" = "Central America and the Caribbean",
  "Seychelles" = "Sub-Saharan Africa", "Dominican Republic" = "Central America and the Caribbean",
  "Albania" = "Europe", "Rwanda" = "Sub-Saharan Africa", "Gabon" = "Sub-Saharan Africa",
  "Canada" = "North America", "Ghana" = "Sub-Saharan Africa",
  "Morocco" = "Middle East and North Africa", "Guinea-Bissau" = "Sub-Saharan Africa",
  "Ireland" = "Europe", "India" = "Asia", "Democratic Republic of the Congo" = "Sub-Saharan Africa",
  "Iceland" = "Europe", "Pakistan" = "Asia", "Laos" = "Asia",
  "Cameroon" = "Sub-Saharan Africa", "Cote d'Ivoire" = "Sub-Saharan Africa",
  "Grenada" = "Central America and the Caribbean", "Maldives" = "Asia",
  "Nepal" = "Asia", "Fiji" = "Australia and Oceania", "Nicaragua" = "Central America and the Caribbean",
  "Jordan" = "Middle East and North Africa", "El Salvador" = "Central America and the Caribbean",
  "Egypt" = "Middle East and North Africa", "Republic of the Congo" = "Sub-Saharan Africa",
  "Angola" = "Sub-Saharan Africa", "Zimbabwe" = "Sub-Saharan Africa", 
  "Nigeria" = "Sub-Saharan Africa"
)
```



#---------------------------- Invalid Region entries are replaced with the correct Regions based on their country

```{r}
fixedCountries <- noNARow
fixedCountries$Region <- ifelse(noNARow$Region %in% validPlaces, noNARow$Region, countryRegionMap[noNARow$Country])

uniqueValues <- unique(fixedCountries[,"Region"])
print(uniqueValues)
```



#------------------------- X is reviewed but we decided to view "None" as the option "other"

```{r}
#Note - None is one of the options
uniqueValues <- unique(fixedCountries[,"X"])
print(uniqueValues)
```

#------------------------- Sales.Channel is checked on what dirty values it has.

```{r}
#Note - Yes is one of the options
uniqueValues <- unique(fixedCountries[,"Sales.Channel"])
print(uniqueValues)
fixedCountries[fixedCountries$Sales.Channel == "", ]
```

```{r}
# Print the unique values in the Sales.Channel column to check
uniqueValue <- unique(fixedCountries[,"Sales.Channel"])
print(uniqueValue)
```


#----------------------- Using predictions we predict if the Sales.Channel based on the country and item type


```{r}

library(dplyr)

# Convert to lowercase to avoid case issues
fixedCountries$Sales.Channel <- tolower(fixedCountries$Sales.Channel)

# Replace only incorrect values ("yes" and "") with NA, keeping "online" and "offline"
fixedCountries$Sales.Channel <- ifelse(fixedCountries$Sales.Channel %in% c("yes", ""), NA, fixedCountries$Sales.Channel)

# Ensure Sales.Channel is a factor
fixedCountries$Sales.Channel <- factor(fixedCountries$Sales.Channel, levels = c("offline", "online"))

# ✅ Check the data before modeling
print(table(fixedCountries$Sales.Channel, useNA = "always"))

# Only run the model if there are enough unique values
train_data <- fixedCountries %>% filter(!is.na(Sales.Channel))

if (length(unique(train_data$Sales.Channel)) > 1) {
    # Train logistic regression model
    model <- glm(Sales.Channel ~ Country + X, data = train_data, family = binomial, na.action = na.exclude)

    # Identify missing values
    missing_indices <- which(is.na(fixedCountries$Sales.Channel))
    fixedCountries$PredictedProb <- predict(model, newdata = fixedCountries, type = "response")

    # Replace missing values based on predicted probability
    fixedCountries$Sales.Channel[missing_indices] <- ifelse(fixedCountries$PredictedProb[missing_indices] > 0.5, "online", "offline")

    # Convert to factor again
    fixedCountries$Sales.Channel <- factor(fixedCountries$Sales.Channel, levels = c("offline", "online"))

    # Remove the PredictedProb column
    fixedCountries <- fixedCountries %>% select(-PredictedProb)
} else {
    warning("Not enough unique levels in Sales.Channel to fit the model. Check data cleaning steps!")
}

# ✅ Check the cleaned data
print(table(fixedCountries$Sales.Channel, useNA = "always"))

```

#------------------------- Order.Date and Ship.Date are compared to check for errors

```{r}

fixedCountries$Order.Date <- as.Date(fixedCountries$Order.Date, format="%m/%d/%Y")
fixedCountries$Ship.Date <- as.Date(fixedCountries$Ship.Date, format="%m/%d/%Y")

all(fixedCountries$Order.Date <= fixedCountries$Ship.Date)


#invalid_rows <- fixedCountries[fixedCountries$Order.Date > fixedCountries$Ship.Date, ]

# Print the problematic rows
#print(invalid_rows)

```


#-------------------------- Region

```{r}
summary(fixedCountries$Region)
uniqueValues <- unique(fixedCountries[,"Region"])
print(uniqueValues)
```




```{r}

product_counts <- fixedCountries %>%
  count(Region) 

ggplot(product_counts, aes(x = reorder(Region, -n), y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
    title = "Frequency of Region",
    x = "Count",
    y = "Region"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-labels

```



#----------------------------------------------- Country

```{r}
summary(fixedCountries$Country)
uniqueValues <- unique(fixedCountries[,"Country"])
print(uniqueValues)
```


```{r}

top_countries <- fixedCountries %>%
  count(Country) %>%
  top_n(10, n) 

ggplot(top_countries, aes(x = reorder(Country, -n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Top 10 Countries by Sales Frequency", x = "Country", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```







#--------------------------------- X

```{r}
summary(fixedCountries$X)
uniqueValues <- unique(fixedCountries[,"X"])
print(uniqueValues)
```

```{r}
product_counts <- fixedCountries %>%
  count(X) 

ggplot(product_counts, aes(x = reorder(X, -n), y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
    title = "Frequency of Product Types (X)",
    x = "Product Type",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-labels
```




#---------------------------------- Sales.Channel


```{r}
summary(fixedCountries$Sales.Channel)
uniqueValues <- unique(fixedCountries[,"Sales.Channel"])
print(uniqueValues)


```
```{r}

channel_counts <- fixedCountries %>%
  count(Sales.Channel) %>%
  arrange(desc(n))  # Sort by frequency (descending)

ggplot(channel_counts, aes(x = Sales.Channel, y = n)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Frequency of Sales by Channel", x = "Sales Channel", y = "Count") +
  theme_minimal()
```


#------------------------------- Order.Priority

```{r}
summary(fixedCountries$Order.Priority)
uniqueValues <- unique(fixedCountries[,"Order.Priority"])
print(uniqueValues)
```

```{r}
channel_counts <- fixedCountries %>%
  count(Order.Priority) %>%
  arrange(desc(n))  # Sort by frequency (descending)

ggplot(channel_counts, aes(x = Order.Priority, y = n)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Frequency of Sales by Channel", x = "Order.Priority", y = "Count") +
  theme_minimal()
```

#---------------------------------- Order.Date

```{r}
summary(fixedCountries$Order.Date)


```

```{r}
ggplot(fixedCountries, aes(x = Order.Date)) +
  geom_line(stat = "count", color = "red", linewidth = .05) +  # Count orders per date
  labs(
    title = "Orders Over Time (Check for Gaps/Spikes)",
    x = "Order Date",
    y = "Number of Orders"
  ) +
  theme_minimal()
```


#---------------------------------- Ship.Date


```{r}
summary(fixedCountries$Ship.Date)

```


```{r}
ggplot(fixedCountries, aes(x = Ship.Date)) +
  geom_line(stat = "count", color = "red", linewidth = .05) +  # Count orders per date
  labs(
    title = "Orders Over Time (Check for Gaps/Spikes)",
    x = "Ship Date",
    y = "Number of Orders"
  ) +
  theme_minimal()

channel_counts <- fixedCountries %>%
  count(Ship.Date) %>%
  arrange(desc(n))  # Sort by frequency (descending)

ggplot(channel_counts, aes(x = Ship.Date, y = n)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Frequency of Sales by Channel", x = "Sales Channel", y = "Count") +
  theme_minimal()


```


#--------------------------------- Order.ID

```{r}
fixedCountries$Order.ID <- as.integer(fixedCountries$Order.ID)
summary(fixedCountries$Order.ID)
```

#--------------------------------- Units.Sold

```{r}
fixedCountries$Units.Sold <- as.integer(fixedCountries$Units.Sold)
summary(fixedCountries$Units.Sold)
```



```{r}
ggplot(fixedCountries, aes(x = Units.Sold)) +
  geom_histogram(bins = 30, fill = "purple", color = "white") +
  labs(title = "Distribution of Units Sold", x = "Units Sold", y = "Frequency") +
  theme_minimal()
```

#---------------------------------- Unit.Cost


```{r}
fixedCountries$Unit.Cost <- as.numeric(fixedCountries$Unit.Cost)
summary(fixedCountries$Unit.Cost)
```


```{r}
# Boxplot for Unit.Price
ggplot(fixedCountries, aes(y = Unit.Price)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Boxplot of Unit Price")

# Scatter Plot: Unit.Price vs. Unit.Cost (should generally be above the diagonal)
ggplot(fixedCountries, aes(x = Unit.Cost, y = Unit.Price)) +
  geom_point(color = "blue") +
  geom_abline(slope = 1, linetype = "dashed", color = "red") +  # Profit=0 line
  labs(title = "Unit Price vs. Unit Cost (Check for Pricing Errors)")
```


#---------------------------------- Total.Cost


```{r}
fixedCountries$Total.Cost <- as.numeric(fixedCountries$Total.Cost)
summary(fixedCountries$Total.Cost)
```


```{r}
# Boxplot
ggplot(fixedCountries, aes(y = Total.Cost)) +
  geom_boxplot(fill = "pink") +
  labs(title = "Boxplot of Total Cost")

# Scatter Plot: Total.Cost vs. Total.Revenue
ggplot(fixedCountries, aes(x = Total.Cost, y = Total.Revenue)) +
  geom_point(color = "darkgreen") +
  geom_abline(slope = 1, linetype = "dashed", color = "red") +  # Break-even line
  labs(title = "Total Cost vs. Total Revenue (Should Mostly Be Below Diagonal)")
```

#------------------------- Total.Revenue


```{r}
fixedCountries$Total.Revenue <- as.numeric(fixedCountries$Total.Revenue)
summary(fixedCountries$Total.Revenue)
```


#--- Using fixedCountries$Total.Profit + fixedCountries$Total.Cost = fixedCountries$Total.Revenue to replace the NA's


```{r}
fixedCountries$Total.Profit <- as.numeric(fixedCountries$Total.Profit)

fixedCountries$Total.Revenue <- ifelse(is.na(fixedCountries$Total.Revenue) & !is.na(fixedCountries$Total.Cost) & !is.na(fixedCountries$Total.Profit),
                          fixedCountries$Total.Profit + fixedCountries$Total.Cost,
                          fixedCountries$Total.Revenue)

summary(fixedCountries$Total.Revenue)
```


```{r}
# Histogram
ggplot(fixedCountries, aes(x = Total.Revenue)) +
  geom_histogram(bins = 30, fill = "steelblue") +
  labs(title = "Distribution of Total Revenue")

# Time Series (if Order.Date exists)
fixedCountries$Order.Date <- as.Date(fixedCountries$Order.Date)  # Convert to date
ggplot(fixedCountries, aes(x = Order.Date, y = Total.Revenue)) +
  geom_line(color = "darkred") +
  labs(title = "Total Revenue Over Time")
```

#------------------------- Unit.Price




```{r}
fixedCountries$Unit.Price <- as.numeric(fixedCountries$Unit.Price)
summary(fixedCountries$Unit.Price)
```

#--- Using fixedCountries$Total.Revenue / fixedCountries$Units.Sold = Unit.Price to replace the NA's

```{r}
fixedCountries$Unit.Price <- ifelse(
  is.na(fixedCountries$Unit.Price) & 
  !is.na(fixedCountries$Total.Revenue) & 
  !is.na(fixedCountries$Units.Sold),
  fixedCountries$Total.Revenue / fixedCountries$Units.Sold,  # Calculate Unit.Price
  fixedCountries$Unit.Price  # Keep original value if not NA
)

summary(fixedCountries$Unit.Price)
```


```{r}
# Boxplot for Unit.Price
ggplot(fixedCountries, aes(y = Unit.Price)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Boxplot of Unit Price")

# Scatter Plot: Unit.Price vs. Unit.Cost (should generally be above the diagonal)
ggplot(fixedCountries, aes(x = Unit.Cost, y = Unit.Price)) +
  geom_point(color = "blue") +
  geom_abline(slope = 1, linetype = "dashed", color = "red") +  # Profit=0 line
  labs(title = "Unit Price vs. Unit Cost (Check for Pricing Errors)")
```


#------------------------- Total.Profit

```{r}
fixedCountries$Total.Profit <- as.numeric(fixedCountries$Total.Profit)
summary(fixedCountries$Total.Profit)
```

#--- Using fixedCountries$Total.Revenue - fixedCountries$Total.Cost = Total.Profit to replace the NA's

```{r}
fixedCountries$Total.Profit <- ifelse(is.na(fixedCountries$Total.Profit) & !is.na(fixedCountries$Total.Revenue) & !is.na(fixedCountries$Total.Cost),
                          fixedCountries$Total.Revenue - fixedCountries$Total.Cost,
                          fixedCountries$Total.Profit)

summary(fixedCountries$Total.Profit)

```


```{r}

# Boxplot (identify extreme values)
ggplot(fixedCountries, aes(y = Total.Profit)) +
  geom_boxplot(fill = "skyblue") +
  labs(title = "Boxplot of Total Profit (Outlier Detection)")

# Histogram (check distribution)
ggplot(fixedCountries, aes(x = Total.Profit)) +
  geom_histogram(bins = 30, fill = "purple") +
  labs(title = "Distribution of Total Profit")

# Density plot (smooth distribution)
ggplot(fixedCountries, aes(x = Total.Profit)) +
  geom_density(fill = "orange", alpha = 0.7) +
  labs(title = "Density of Total Profit")
```


#-------------------------




