---
title: "Assignment02"
output: html_document
date: "2025-03-28"
---

## Project Overview

A new lifestyle and retail brand has approached you to help them assess data related to buying patterns
in different countries. They are in an expansion mode and would like to present some findings to their
investment board regarding the pros and cons related to some key products that they carry. They are
also interested also in understanding perspectives from travelers in specified countries to enhance their
understanding of how to approach a key segment of their market, vacationers

```{r setup}
#install.packages(c("dplyr", "tidyr", "ggplot2", "plotly", "tm"))
#install.packages("corrplot")
#install.packages(c("SnowballC","syuzhet","wordcloud","wordcloud2","RColorBrewer"))


library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(tm)
library(corrplot)
library(SnowballC)
library(syuzhet)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)


```
## Part 1

First we will load in traveler data, and create dataframes for three Caribbean countries: Jamaica, The Bahamas and Trinidad and Tobego
```{r load data}

#Clean up data
rm(list=ls())
options(scipen=9999)

#Load in traveler data
tdata <- read.csv("traveler_reviews.csv")
head(tdata, 20)

#Create subsets of the data for Jamaica, The Bahamas and Trinidad
jam <- subset(tdata, Country == "Jamaica")
head(jam)
baha <- subset(tdata, Country == "The Bahamas")
head(baha)
trini <- subset(tdata, Country == "Trinidad and Tobago")
head(trini)
```
## Wordclouds

# The Bahamas
```{r Bahamas wordcloud}
#Get reviews for wordclouds
baha_rev <- baha$ReviewText

#Convert to corpus and convert to lowercase
btext <- Corpus(VectorSource(baha_rev))
btext <- tm_map(btext, content_transformer(tolower))


#Remove stopwords
btext <- tm_map(btext, removeWords, c(stopwords("english"), "and", "or","hotel"))


#Create wordcloud
btext.dtm <- TermDocumentMatrix(btext)

btext.matrix <- as.matrix(btext.dtm)

bwords <- sort(rowSums(btext.matrix), decreasing=TRUE)

btext.df <- data.frame(word = names(bwords),freq=bwords)
wordcloud2(data=btext.df, size = 0.20, shape="ellipse")

```
# Jamaica 

```{r Jamaica wordcloud}
#Get reviews for wordclouds
jam_rev <- jam$ReviewText

#Convert to corpus and convert to lowercase
jtext <- Corpus(VectorSource(jam_rev))
jtext <- tm_map(jtext, content_transformer(tolower))

#Remove stopwords
jtext <- tm_map(jtext, removeWords, c(stopwords("english"), "and", "or","hotel"))

#Create wordcloud
jtext.dtm <- TermDocumentMatrix(jtext)

jtext.matrix <- as.matrix(jtext.dtm)

jwords <- sort(rowSums(jtext.matrix), decreasing=TRUE)

jtext.df <- data.frame(word = names(jwords),freq=jwords)
wordcloud2(data=jtext.df, size = 0.20, shape="ellipse")
```
# Trinidad
```{r Trinidad}
#Get reviews for wordclouds
trini_rev <- trini$ReviewText
head(trini_rev)

#Convert to corpus and convert to lowercase
ttext <- Corpus(VectorSource(trini_rev))
ttext <- tm_map(ttext, content_transformer(tolower))

#Remove stopwords
ttext <- tm_map(ttext, removeWords, c(stopwords("english"), "and", "or","hotel"))

#Create wordcloud
ttext.dtm <- TermDocumentMatrix(ttext)

ttext.matrix <- as.matrix(ttext.dtm)

twords <- sort(rowSums(ttext.matrix), decreasing=TRUE)

ttext.df <- data.frame(word = names(twords),freq=twords)
wordcloud2(data=ttext.df, size = 0.20, shape="ellipse")
```
## Wordcloud Comparisons
Based on the word clouds, it is apparent that most reviews for Jamaica and Trinidad and Tobego are about the hotel room, since the largest word in their wordclouds is "room". On the other hand, most reviews in the Bahamas discuss "breakfast"

## Sentiment Analysis for each Country

# Bahamas

```{r bahamas sentiment }
#Get sentiments and add to table
bsent <- get_nrc_sentiment(baha$ReviewText)
bvibes <- cbind(baha, bsent)
head(bvibes)

#Determine dominant sentiment
count <- colSums(bsent)
most <- names(count)[which.max(count)]
cat("The most common sentiment in The Bahamas is", most)

```

#Jamaica

```{r jamaica sentiment }
#Get sentiments and add to table
jsent <- get_nrc_sentiment(jam$ReviewText)
jvibes <- cbind(jam, jsent)
head(jvibes)

#Determine dominant sentiment
count <- colSums(jsent)
most <- names(count)[which.max(count)]
cat("The most common sentiment in Jamaica is", most)

```

#Trinidad and Tobego

```{r trinidad sentiment }
#Get sentiments and add to table
tsent <- get_nrc_sentiment(trini$ReviewText)
tvibes <- cbind(trini, tsent)
head(tvibes)

#Determine dominant sentiment
count <- colSums(tsent)
most <- names(count)[which.max(count)]
cat("The most common sentiment in Trinidad and Tobago is", most)

```

## Analysis


# What is the average amount spent on Personal Care vs. Cosmetics in each region?

```{r a}

#Clean up data
rm(list=ls())
options(scipen=9999)

sales.df = read.csv("sales_data03.csv", stringsAsFactors = FALSE)


#Convert Revenue to numeric
sales.df$Total.Revenue <- as.numeric(sales.df$Total.Revenue)

#Create dataframes for Personal Care and Cosmetics
caredata <- subset(sales.df, X == "Personal Care")


cosmeticsdata <- subset(sales.df, X == "Cosmetics")


careavg <- mean(caredata$Total.Revenue, na.rm = TRUE)
cosmeticsavg <- mean(cosmeticsdata$Total.Revenue, na.rm = TRUE)

cat("The average money spent on Personal care is", careavg, "while the average spent on Cosmetics is", cosmeticsavg)
```
# In which country do people use the most cosmetics? 

```{r b}
#Convert Revenue to numeric
cosmeticsdata$Units.Sold <- as.numeric(cosmeticsdata$Units.Sold)

#Get sums the units sold in each country
cosmeticsum <- aggregate(Units.Sold ~ Country, data = cosmeticsdata, FUN = sum, na.rm = TRUE)

#Gets the data for the country with the maximum units sold
cosmax <- cosmeticsum[which.max(cosmeticsum$Units.Sold), ]
cat("People use the most cosmetics in", cosmax$Country)
```
# In 2016, which product had the highest and lowest quantity sold across each region?

```{r c}
#Set datatypes
sales.df$Order.Date <- as.Date(sales.df$Order.Date, format = "%d/%m/%Y")
sales.df$Units.Sold <- as.numeric(sales.df$Units.Sold)

#Create subset of data from 2016
data2016 <- subset(sales.df, format(Order.Date, "%Y") == "2016")

#Find total per region and per product
total <- aggregate(Units.Sold ~ Region + X, data = data2016, sum, na.rm = TRUE)

#Split data by regions
byregions <- split(total, total$Region)

#Create a dataframe with the maximums and minimums found from byregions

hilo <- lapply(byregions, function(df) {
  data.frame(
    Region = unique(df$Region),
    Highest.Sold = df$X[which.max(df$Units.Sold)],
    Lowest.Sold = df$X[which.min(df$Units.Sold)]
  )
})

hilo <- do.call(rbind, hilo)
print(hilo)

```
# What is the average profit for each item by region? 

```{r d}
#Set datatypes
sales.df$Total.Profit <- as.numeric(sales.df$Total.Profit)

aggregate(Total.Profit ~ Region + X, data = sales.df, FUN = mean, na.rm = TRUE)
```
# In which of the seasons (Spring, Summer, Autumn, Winter) do people spend the most on:  
Cosmetics
Personal Care
```{r e}
#Note that we will use the same seasons for all regions, and assume typical Northern Hemisphere Seasons where, generally
#Winter is December, January and February
#Spring is March, April and May
#Summer is June, July and August
#Autumn is September, October, November

#Set datatypes
sales.df$Month <- as.numeric(format(sales.df$Order.Date, "%m"))

#Create bins labelled as seasons, and assign months to seasons
sales.df$Season <- cut(sales.df$Month,
                       breaks = c(0, 2, 5, 8, 11, 12),
                       labels = c("Winter", "Spring", "Summer", "Autumn", "Winter"),
                       include.lowest = TRUE, right = TRUE)

#Create an aggregate of revenue by season
seasonsales <- aggregate(Total.Revenue ~ Season + X, data = sales.df[sales.df$X %in% c("Cosmetics", "Personal Care"), ], sum, na.rm = TRUE)


#Set datatypes
seasonsales$Season <- as.character(seasonsales$Season)

#Get maximums
maxcos <- seasonsales$Season[which.max(seasonsales$Total.Revenue[seasonsales$X == "Cosmetics"])]
maxcare <- seasonsales$Season[which.max(seasonsales$Total.Revenue[seasonsales$X == "Personal Care"])]

cat("Highest spending on cosmetics was in", maxcos, "and highest spending on personal care was in", maxcare)
```
# Find the monthly average units sold for Personal Care, Cosmetics and Beverages and plot a series graph for ONE year . Discuss the resulting visualization.
```{r f}
#Remove days
data2016$YearMonth <- format(data2016$Order.Date, "%Y-%m")

#Create subset with only the needed data
data2016filter<- data2016[data2016$X %in% c("Personal Care", "Cosmetics", "Beverages"), ]

#Find monthly average
monthlyavg <- aggregate(Units.Sold ~ YearMonth + X, data = data2016filter, FUN = mean, na.rm = TRUE)

#Plot line graph
ggplot(monthlyavg, aes(x = YearMonth, y = Units.Sold, color = X, group = X)) + geom_line() + labs(title = "Monthly Average Units Sold", x = "Month", y = "Units Sold") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Determine correlation between Total.Profit and the other features (include visualizations) in the dataset.

Units Sold and Total Profit are x correlated
Unit Price and Total Profit are x correlated
Unit Cost and Total Profit are x correlated
Total Revenue and Total Profit are x correlated
Total Cost and Total Profit are x correlated
```{r g}
#Set datatypes
sales.df <- sales.df %>% select(Total.Profit, Units.Sold, Unit.Price, Unit.Cost, Total.Revenue, Total.Cost) %>%mutate(across(everything(), as.numeric))

#Create correlation matrix
num_cols <- sapply(sales.df, is.numeric)
cor_matrix <- cor(sales.df[, num_cols], use = "complete.obs")
profit_corr <- cor_matrix["Total.Profit", , drop = FALSE]

corrplot(profit_corr, method = "color")

#Check correlations
cor(sales.df$Units.Sold,sales.df$Total.Profit)
cor(sales.df$Unit.Price,sales.df$Total.Profit)
cor(sales.df$Unit.Cost,sales.df$Total.Profit)
cor(sales.df$Total.Revenue,sales.df$Total.Profit)
cor(sales.df$Total.Cost,sales.df$Total.Profit)
```