---
title: "Assignment02"
output: html_document
date: "2025-03-28"
---

## Project Overview

A new lifestyle and retail brand has approached you to help them assess data related to buying patterns
in different countries. They are in an expansion mode and would like to present some findings to their
investment board regarding the pros and cons related to some key products that they carry. They are
also interested also in understanding perspectives from travelers in specified countries to enhance their
understanding of how to approach a key segment of their market, vacationers

```{r setup}
#install.packages(c("dplyr", "tidyr", "ggplot2", "plotly", "tm"))
#install.packages("corrplot")
install.packages(c("SnowballC","syuzhet","wordcloud","wordcloud2","RColorBrewer"))


library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(tm)
library(corrplot)
library(SnowballC)
library(syuzhet)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)


```
## Part 1

First we will load in traveler data, and create dataframes for three Caribbean countries: Jamaica, The Bahamas and Trinidad and Tobego
```{r load data}

#Clean up data
rm(list=ls())
options(scipen=9999)

#Load in traveler data
tdata <- read.csv("traveler_reviews.csv")
head(tdata, 20)

#Create subsets of the data for Jamaica, The Bahamas and Trinidad
jam <- subset(tdata, Country == "Jamaica")
head(jam)
baha <- subset(tdata, Country == "The Bahamas")
head(baha)
trini <- subset(tdata, Country == "Trinidad and Tobago")
head(trini)
```
## Wordclouds

# The Bahamas
```{r Bahamas wordcloud}
#Get reviews for wordclouds
baha_rev <- baha$ReviewText

#Convert to corpus and convert to lowercase
btext <- Corpus(VectorSource(baha_rev))
btext <- tm_map(btext, content_transformer(tolower))


#Remove stopwords
btext <- tm_map(btext, removeWords, c(stopwords("english"), "and", "or","hotel"))


#Create wordcloud
btext.dtm <- TermDocumentMatrix(btext)

btext.matrix <- as.matrix(btext.dtm)

bwords <- sort(rowSums(btext.matrix), decreasing=TRUE)

btext.df <- data.frame(word = names(bwords),freq=bwords)
wordcloud2(data=btext.df, size = 0.20, shape="ellipse")

```
# Jamaica 

```{r Jamaica wordcloud}
#Get reviews for wordclouds
jam_rev <- jam$ReviewText

#Convert to corpus and convert to lowercase
jtext <- Corpus(VectorSource(jam_rev))
jtext <- tm_map(jtext, content_transformer(tolower))

#Remove stopwords
jtext <- tm_map(jtext, removeWords, c(stopwords("english"), "and", "or","hotel"))

#Create wordcloud
jtext.dtm <- TermDocumentMatrix(jtext)

jtext.matrix <- as.matrix(jtext.dtm)

jwords <- sort(rowSums(jtext.matrix), decreasing=TRUE)

jtext.df <- data.frame(word = names(jwords),freq=jwords)
wordcloud2(data=jtext.df, size = 0.20, shape="ellipse")
```
# Trinidad
```{r Trinidad}
#Get reviews for wordclouds
trini_rev <- trini$ReviewText
head(trini_rev)

#Convert to corpus and convert to lowercase
ttext <- Corpus(VectorSource(trini_rev))
ttext <- tm_map(ttext, content_transformer(tolower))

#Remove stopwords
ttext <- tm_map(ttext, removeWords, c(stopwords("english"), "and", "or","hotel"))

#Create wordcloud
ttext.dtm <- TermDocumentMatrix(ttext)

ttext.matrix <- as.matrix(ttext.dtm)

twords <- sort(rowSums(ttext.matrix), decreasing=TRUE)

ttext.df <- data.frame(word = names(twords),freq=twords)
wordcloud2(data=ttext.df, size = 0.20, shape="ellipse")
```

```{r}
the_tweets.df = read.csv("sales_data03.csv", stringsAsFactors = FALSE)
```



