---
title: "Assignment02"
output: html_document
date: "2025-03-28"
---

## Project Overview

A new lifestyle and retail brand has approached you to help them assess data related to buying patterns
in different countries. They are in an expansion mode and would like to present some findings to their
investment board regarding the pros and cons related to some key products that they carry. They are
also interested also in understanding perspectives from travelers in specified countries to enhance their
understanding of how to approach a key segment of their market, vacationers

```{r setup}
#install.packages(c("dplyr", "tidyr", "ggplot2", "plotly", "tm"))
#install.packages("corrplot")
install.packages(c("SnowballC","syuzhet","wordcloud","wordcloud2","RColorBrewer"))


library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(tm)
library(corrplot)
library(SnowballC)
library(syuzhet)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)


```
## Part 1
```{r load data}

#Clean up data
rm(list=ls())
options(scipen=9999)

#Load in traveler data
tdata <- read.csv("traveler_reviews.csv")
head(tdata, 20)

#Create subsets of the data for Jamaica, The Bahamas and Trinidad
jam <- subset(tdata, Country == "Jamaica")
head(jam)
baha <- subset(tdata, Country == "The Bahamas")
head(baha)
trini <- subset(tdata, Country == "Trinidad and Tobago")
head(trini)
```





```{r}
the_tweets.df = read.csv("sales_data03.csv", stringsAsFactors = FALSE)
```



